A Physically Motivated Framework to Compare Merger Timescales\\ of Isolated Low- and High-Mass Galaxy Pairs Across Cosmic Time

### 2024-03-29 paper/main.tex
complete discussion and conclusion, and half finished intro!!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction} \label{sec:intro}
% Outline draft - 
% 1 - Galaxy mergers are an importance feature of LCDM theory and are critical for the formation and evolution of galaxies across time. 
%     - An important feature of galaxy mergers is the merger timescale...

% 2 - Merger timescales are important to calibrate 
%     - They are crucial for understanding the observations of galaxy morphologies and understanding heirarchical evolution, (think galaxy start bursts, formation of gas bridges maybe, fueling agn, BH merger rates, etc)
    %- Also important for understanding the few well modelled orbits that we do have, particularly in the LG (might come way later)

% 3 - Merger timescales are also used to get merger rates, both observationally and theoretically, of close pairs.
%     - How are close pairs defined? 
%     - How do merger timescales enter merger rate calculations?
%     - What is important about the merger rates that are used? 
%     - Maybe need to make distinction here that there are also observability timescales, but they are going to be basically useless for TNG100 studies (see below), but also are low mass pairs have a lot of gas and look super weird.

Galaxy merger rates are a test of LCDM theory. In the era of JWST, these tests can push to higher redshifts and lower mass scales.  At high redshift and at lower masses, even isolated galaxies are highly disturbed, making pair fractions critical for merger rate studies \cite{Martin2018, RG2019}. However, there is no framework for comparing merger rates from pair fractions for low mass versus high mass galaxies and tensions exist in the literature at low masses (e.g. Vicente paper figure). i.e. There is a tension at low mass% vicente's work is done consistently, but we're basically showing that if you want to put the observational stuff on there too, you can't just willy nilly use any old pair criteria

This study seeks to establish a framework to enable comparisons of merger rates across mass scales and redshift so that merger rates of low mass galaxies can be used as tests of LCDM theory. These results can also be used to interpret observations of galaxy morphologies and their hierarchical evolution. , (think galaxy start bursts, formation of gas bridges maybe, fueling agn, BH merger rates, etc)

@@

% 5 - Merger timescales have been studied a lot (mostly in obs and theory of high mass pairs)
%     - Lotz 2011
%     - RG2015
%     - Snyder (maybe?) 
%     - Jiang 2008 \& BK 2008 for all things about the "Chandra" timescale for merging galaxies which scales with rc/Vc and is thus prop to the age of the universe (eq. 4\&5). Also quantified the circularity of orbits as a function of mass ratio $\leftarrow$ might be good to mention
%     - Jiang 2014

@@

9 - Here, we aim to build a framework to understand the pair selection criteria that are used to select samples, and the merger timescales that result from various selections.  we look at the orbits of the pairs from our previous work, and determine the merger timescale for low and high mass pairs as a function of pair separation and redshift. In sec. 1... etc.

%%% OUTLINE
\section{Methodology} \label{sec:methods}
%recap last paper and mention which data sample we narrow down to (low mass major pairs only) -- give examples of orbits
We utilize the group catalogs, produced by the \subfind{} algorithm~\citep{Springel2001b,Dolag2009}, and the merger tree catalogs, generated by the \sublink{} algorithm~\citep{RG2015}, from the highest resolution run of the IllustrisTNG simulation TNG100-1 (hereafter TNG100) to study the merger timescales of galaxy pairs from $z=0-6$.
The catalogs consist of a set of 100 snapshots, ranging from $z\sim20$ (snapshot 0) to $z=0$ (snapshot 99).

@@

We require that subhalos constituting a pair meet a minimum subhalo mass 
% GB added Mhalo
($\Mhalo$)
criteria of 
\begin{equation*}
    \mbox{\textbf{minimum subhalo mass:}}\,
    \Mhalo > 1\times10^{9}\Msun.
\end{equation*}
at the snapshot of consideration. 
For TNG100, this ensures that subhalos are resolved into over 
$\sim$100 particles, enough to robustly identify gravitationally bound subhalos in the \subfind{} and \sublink{} catalogs.
For each subhalo in the FoF group that passes the minimum subhalo mass criteria, we utilize the \sublink{} catalogs to find the peak halo mass of each subhalo \citep{RG2015}.

% GB added 
As in \chambe{}, 
stellar masses are assigned to each subhalo in the FoF group using the abundance matching prescription of \citet{Moster2013}.
The peak halo mass and current redshift are used to calculate the stellar mass of each subhalo via the abundance matching prescription  $\ms{}=f(\Mpeak,z)$, where $z$ is the redshift of consideration.

@@

\subsection{Number of Pairs}\label{sec:pairprops-num}
We calculate the number of pairs by selecting all pairs at a given redshift that have not merged.
% The number of orbits that result in mergers at a later redshift are selected in the same way, excluding the non-merging orbits.
%For example, the total number of orbits at $z=2$ includes the orbits of pairs that passed the selection criteria at \textit{any} redshift, provided they achieved first infall at $z\geq2$ and have a merger redshift $z<2$. 
A single, non-merging pair with first infall at $z=3$ will thus contribute to the number of pairs at all redshifts from $z=0-3$.

@@

\subsection{Merger Fraction}\label{sec:pairprops-frac}
We calculate the merger fraction by dividing the number of pairs that merge (dashed lines in Fig.~\ref{fig:numorbits}) by the number of total unique orbits (solid lines) at a given redshift, including all pairs that do and do not merge. 
As before, an orbit with first infall at $z=2$ and merger at $z=1$ will be included in the merger fraction calculation for redshifts $z=1-2$.

@@

% difference/similiarity between low and high
The merger fraction of low-mass and high-mass pairs is the same from $z=0-0.5$ and $z>2.5$. 
Between $z=0.5-2.5$, the high-mass merger fraction is higher than the low-mass merger fraction, and the knee of the merger fraction occurs at a lower redshift.
This is likely because the number of low-mass pairs continues to increase before turning over and beginning to decrease at $z\sim2$, while the number of high-mass pairs continues increasing to $z\sim1$. 
At these redshifts, the number of merging pairs begins declining rapidly. 
% When the number of pairs stops increasing and beings to decrease, pairs that will not have time to merge before $z=0$ are being added, while many merging pairs are being removed from the sample, thus decreasing the merger fraction. 
Since the turnover for the number of high-mass pairs occurs at a lower redshift, the knee of the merger fraction is likewise shifted to lower redshifts.
% move the above to discussion and end with something shorter: 
% This COULD be that the timescales. This will answer this with this study and discuss this in detail in Disc.~\ref.

@@

\section{Results: The Mass and Redshift Dependence of Merger Timescales}
    Using our sample of isolated low-mass and high-mass pairs in the TNG100 simulation, we calculate the merger timescales, or time until merger, for all of the merging pairs in our catalog. 
    The merger timescale of a pair is defined to be the amount of time that elapses in the simulation between the redshift at which pairs are selected and the merger redshift.\footnote{Since our orbits are only defined after infall, the merger timescales will likewise only be calculated for pairs that are post-infall at the given separation.}
    Only the merging pairs will be considered for the remainder of the analysis.

@@

\subsection{Separation Dependence of Merger Timescales}\label{sec:results-timevsep}
    % how we calculate for this section
    We calculate the time until merger as a function of separation for low-mass and high-mass pairs at a variety of redshifts from $z=0.5-6$. 
    % Pairs with small separations may be undergoing a pericentric passage, or very close to merging if on a radial orbit.
    Binning the pairs by separation, we can study how the merger timescale changes for pairs selected at different points in their orbits.

@@

    We find that the time until merger for low-mass and high-mass pairs is positively correlated with the separation of the pair at each redshift. 
    However, the slope of this correlation decreases at higher redshifts. 
    % For the same physical separation, the merger timescale is lower than for low-mass 
    % This does that the overall merger timescales are lower for high-mass systems. 
    % This is discussed in detail in the next section.
    Pairs at $z=4$ with separations between $50-100\,\kpc$ tend to merge in less time than pairs with the same separations at $z=0.5$.

    In addition, the spread of times until merger and separations is smaller at higher redshift, for both low- and high-mass pairs. 
    The spread of merger timescales for low-mass pairs at $z=1$ goes from $0-8\,\Gyr$, while pairs at $z=4$ have merger timescales between $0-3\,\Gyr$. 
    Likewise, the spread of separations at $z=1$ is $0-200\kpc$, but at $z=4$ the spread is smaller, between $10-125\kpc$. 
    The high-mass pairs also have a larger spread in the distribution of merger timescales and separations at low redshift than at higher redshift.
    This is due to the growth of halos over time, such that more pairs are identified at low redshifts where the virial radius of FoF groups is larger.

@@

    \begin{figure*}[htb]
        \centering
        \includegraphics[width=\textwidth]{plots/bet-on-it/3_time_til_merger.png}
        \caption{(Top) The median time until merger as a function of redshift for low-mass (green) and high-mass (pink) pairs with 3D physical separations greater than $10\,\kpc$ and less than $(50,70,100,150,200,\mbox{and }300)\,\kpc$ from left to right. 
        (Bottom) The median time until merger as a function of redshift for pairs with physical separations greater than $10\,\kpc$ and less than $(0.25, 0.5, 1, 1.5, 2,\mbox{and }2.5)\,\Rvir$ from left to right. 
        % 
        The time until merger increases from $z=4$ to $z\sim0.5-0.75$, at which point the time until merger decreases to zero since all low redshift mergers must have short merger timescales to merge before the end of the simulation at $z=0$. 
        }
        \label{fig:timescales-sep}
    \end{figure*}

@@

        % redshift evolution
        The median time until merger is $\sim0.5-0.7\,\Gyr$ at $z\sim6$, then rises to a peak of $2.3-2.4\,\Gyr$ at $z\sim0.6$, then decreases to 0 at $z=0$.
        The abrupt decrease is artificial rather than a true physical feature. 
        As $z\to0$, there is an increasingly small fraction of pairs that merge before $z=0$, as seen by the dashed lines in Fig~\ref{fig:numorbits}. 
        These mergers must then proceed on timescales less than the elapsed time difference between a given redshift and $z=0$. These mergers must then proceed on decreasing timescales by definition, as they were selected as pairs that must merge prior to $z=0$ (i.e., a pair chosen by the selection criteria given in Section XYZ at z=0.2 must merge by $z=0$, giving a maximal merger timescale equivalent XYZ Gyr whereas a pair chosen at z=6 can have a maximal merger timescale of XYZ Gyr.). 
        This is shown by the dotted black line on the left of Fig.~\ref{fig:timescales}, which shows the lookback time of the simulation as a function of redshift.\footnote{The lookback time at a given redshift $z_n$ is equivalent to the time elapsed between $z_n$ and $z=0$. Thus, merging pairs cannot have merger timescales larger than the lookback time of the simulation at a given redshift.} % this could go in the methods section, if we think it's more necessary there. I don't mind it here for more detail

        In principle, merger timescales\footnote{Specifically, the elapsed time between crossing into the virial radius of the primary subhalo and coalescence with the primary.} for a body moving through a homogeneous field of collision-less matter can be analytically derived from the Chandrasekhar formula for dynamical friction~\citep{Binney2008}. 
        Departures from the idealized case introduce perturbations to that solution, the validity of which has been tested in cosmological hydrodynamic and N-body simulations~\citep{Jiang2008, Boylan_kolchin2008}.

@@

        In Fig.~\ref{fig:timescales}, the black dashed line represents a merger timescale that scales with $1/H(z)$,
        \footnote{$H(z)$ is calculated using the same cosmology as the TNG100 simulation. Specifically, $\Omega_M=0.31$ and $\Omega_{\Lambda}=0.69$.}
        and which is consistent with our findings for redshifts between $z=1.5-6$.

    \subsubsection{Physical Separation Selected Pairs }
    \label{sec:results-phys}
        Both separation selected samples have a lower separation criteria of $10\,\kpc$, to limit the impact of subhalos becoming indistinguishable in the \subfind{} catalogs.
        This lower separation criteria is also commonly applied to observationally selected pairs in studies of merger fractions and merger rates (see Sec.~\ref{sec:discussion} for further detail).

@@

% ##################
% Discussion section
\section{Discussion} \label{sec:discussion}
    We have calculated the merger timescales, or time until merger, for a sample of physically collocated low-mass and high-mass pairs from $z=0-6$.
    In this section, we will explore the implications and broader impacts of our findings, and put our study in context with previous work. 
    %In Sec.~\ref{}, etc...

    In Sec.~\ref{sec:disc-nonmergers}, we will discuss the non-merger portion of our orbit sample and the difference between low- and high-mass merger fractions.
    In Sec.~\ref{sec:disc-previous}, we discuss our work in the context of other studies that have sought to reconcile xx xmerger rates and merger timescales via a change in pair selection criteria. 
    In Sec.~\ref{sec:disc-suggestions}, we discuss the implications of our work for future observational and theoretical merger fraction and merger rate studies and provide suggestions for a self-consistent framework moving forward.
    Finally, in Sec.~\ref{sec:disc-phys}, we briefly discuss the underlying physical mechanisms implied by our results and the functional form of the merger timescale.

    \subsection{Non-merger population}\label{sec:disc-nonmergers}
    In our catalog of 22,213 low-mass pairs and 3,029 high-mass pairs, only 16.66\% (3700) of low-mass pairs and 29.71\% (903) of high-mass pairs do not merge by $z=0$.
    In Sec.~\ref{sec:pairprops-frac}, we found that the merger fraction of low-mass and high-mass pairs at $z>3$ is $>0.95$. 
    This changes as you move to lower redshifts, where the merger fraction decreases substantially. 
    While the decline in the merger fraction at $z\lesssim1$ is primarily due to the limited time that a pair will have to merge before the end of the simulation, we also find that between $z=\sim0.6-2.5$ the fraction of low-mass pairs is lower than that of high-mass pairs.

    One potential explanation for this behavior could be that low-mass pairs have shorter merger timescales than high mass pairs, and thus merge at earlier times in the simulations, thus leading to a decrease in the merger fraction at higher redshifts. 
    However, through our analysis, we have shown that the merger timescales of our full low-mass and high-mass pair sample evolve in the same way from $z=0-6$. 
    This means that the difference in the merger fraction between $z=0.5-2.5$ is not a result of different merger timescales.

    We suspect that the cause of the difference between low- and high-mass merger fractions is not a difference in the merging population itself, but rather in the non-merger population and its evolution over time. 
    As we did not distinguish our non-merging sample into "likely mergers" and "fly-by" interactions for the purposes of the study, we cannot say anything about how these two populations individually contribute to the difference in the merger fraction.
    However, one possible explanation could be that a larger fraction of infalling low-mass pairs at lower redshift result in fly-by interactions. 
    This could be the case if the velocity at which secondary subhalo enter the primary's FoF group gets increasingly large compared to the virial velocity\footnote{The virial velocity is the value of the circular velocity at the virial radius.}, resulting in fewer bound systems. 
    We leave this as a focus of future work.

@@

    \subsection{Steps Towards Standardized Pair Selection Criteria}~\label{sec:disc-previous}
    Comparisons between studies of pair fractions, merger fractions, and merger rates across cosmic time and different mass scales are often made challenging by the use of different sets of selection criteria in each study. 
    On the observational side, pair selection criteria are often set by the observational parameters of the survey itself. 
    For example, limiting magnitudes and completeness limits can enforce a specific new range of stellar mass ratios to be considered (for example $1-1:6$ instead of $1:10$), while avoiding colliding/overlapping fibers on a spectroscopic survey might lead to very specific separation criteria.

    On the theoretical side, it can be straightforward to adopt a number of different pair selection criteria for comparisons to specific observational studies with identical criteria. 
    However, this means that theoretical studies typically restrict themselves to a subset of pair selection criteria specific to their target comparison studies.

    A variety of studies in recent years aim to standardize pair selection criteria for future studies, in particular to permit a more fair comparison between theory and observations. 
    One such study is~\citet{Ventou2017}, which created mock catalogs from the original Illustris-1 simulation to analyze the merger probability of galaxy pairs as a function of projected separation and projected relative velocity. 
    They find that 
    to provide a new set of selection criteria to define close pairs. 
    One

    Merger rates/merger fractions in illustris, close pair fractions and merger timescales, etc. 
    As in our previous work on pair fractions in TNG100, we are unable to make meaningful comparisons with previous observational studies of merger fraction and merger rates, as our study is completed using true physical separations from the simulation. 
    However, 
    % In~\cite{RG2015}, they quantify the galaxy major merger rate as a function of the stellar mass of the descendent subhalo. 
    % It is unclear, however, if the major merger rate for galaxies with stellar mass $\ms<10^{10}\Msun$ is actually discrepant between simulations and observations, or if the observational sam

    NOT a lit review of merger topics etc. 
        \begin{itemize}
            \item Have they studied merger timescales as a function of mass and redshift? 
            \item What is unique about our study?
            \item Are comparisons with previous works fair, or are they not really comparable?
            \item specific papers: Ventou 2017, Hopkins 2007, O'Leary 20
            \item only look at merger rates on the low mass end ? - garreth 
            \item studies that redefine pair fraction criteria 
            \item What other studies have looked at merger fractions/timescales in hydro simulations, specifically? 
            \item Have they studied merger fractions as a function of mass and redshift? What did they find?
            \item Were low-mass pairs expected to have such high merger rates in isolated environments? 
        \end{itemize}

    \subsection{Implications for Future Observational and Theoretical Studies of Pairs}~\label{sec:disc-suggestions}
    In Sec.~\ref{sec:results-timevredshift}, we find that the merger timescales of isolated low-mass and high-mass pairs evolve nearly identically between $z=0-6$, rising from a median of $\sim0.8Gyr$ at $z=6$ to $2.1-2.4\Gyr$ at $z\sim0.6$.
    From $z=0$ to $z=0.6$, the merger timescale rapidly declines, since the only mergers at low redshift that can proceed before the end of the simulation at $z=0$ are those with very short timescales.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction} \label{sec:intro}
Galaxy merger rates provide an important pathway for tests of hierarchical assembly from$\Lambda$CDM theory, and are critical for understanding the formation and evolution of galaxies across time~\citep[e.g.][]{Stewart2009, Hopkins2010b,RG2015}. 
In the era of future surveys with JWST, Rubin, and Roman, these tests will be extended to higher redshifts and lower mass scales than were previously accessible~\citep{Gardner2006,Spergel2015,Robertson2019a,Robertson2019b,Behroozi2020,Martin2022}.

At higher redshift ($z>\kc{xx}$) and at lower masses ($\ms{} \lesssim10^{9.5}$), even isolated galaxies are highly disturbed~\citep{Wuyts2013,Martin2018,RG2019,Martin2021}, making close pair fractions a critical alternative to morphological signatures for merger rate studies.
However, there is currently no framework for comparing merger rates for low-mass and high-mass galaxies as a function of time, and tensions between predictions from theory and observational studies exist in the literature at low masses~\citep[see e.g. Fig.~10 in][]{RG2015}.

This study seeks to establish a framework to enable comparisons of merger rates across mass scales and redshift so that merger rates of low-mass galaxies can be used as tests of $\Lambda$CDM theory. 
These results can also be used to interpret observations of galaxy morphologies and their hierarchical evolution, including the role of mergers in triggering starbursts, fueling active galactic nuclei (AGN), facilitating the formation of tidal features, etc., in low-mass galaxies~\citep[e.g.][]{Stierwalt2015,Pearson2016,Privon2017,Kristensen2021,Martin2021,Luber2022,Martin2022,GuzmanOrtega2023,ByrneMamahit2024,KadoFong2024}.

@@

@@

9 - Here, we aim to build a framework to understand the pair selection criteria that are used to select samples, and the merger timescales that result from various selections.  we look at the orbits of the pairs from our previous work, and determine the merger timescale for low and high mass pairs as a function of pair separation and redshift.

%%% OUTLINE
In Sec.~\ref{sec:methods}, we detail our selection criteria for isolated low-mass and high-mass orbits in the TNG100 simulation. 
We examine the evolution of the number of pairs and merger fraction of our sample as a function of redshift in Sec.~\ref{sec:pairprops}.
In Sec.~\ref{sec:results}, we present our findings on the separation and redshift dependence of the merger timescales of low-mass and high-mass pairs, and study the impact of physical and scaled separation criteria for pair selection. 
Finally, we discuss the implications of this work and provide suggestions for a self-consistent way of studying pairs across different redshifts and mass scales in Sec.~\ref{sec:discussion}, and present our final conclusions in Sec.~\ref{sec:conclusions}.

\section{Methodology} \label{sec:methods}
%recap last paper and mention which data sample we narrow down to (low mass major pairs only) -- give examples of orbits
We utilize the group catalogs, produced by the \subfind{} algorithm~\citep{Springel2001b,Dolag2009}, and the merger tree catalogs, generated by the \sublink{} algorithm~\citep{RG2015}, from the highest resolution run of the IllustrisTNG simulation TNG100-1 (hereafter TNG100) to study the merger timescales of galaxy pairs from $z=0-6$.
The catalogs consist of a set of 100 snapshots, ranging from $z\sim20$ (snapshot 0) to $z=0$ (snapshot 99).

@@

We require that subhalos constituting a pair meet a minimum subhalo mass ($\Mhalo$) criteria of 
\begin{equation*}
    \mbox{\textbf{minimum subhalo mass:}}\,
    \Mhalo > 1\times10^{9}\Msun.
\end{equation*}
at the snapshot of consideration. 
For TNG100, this ensures that subhalos are resolved into over 
$\sim$100 particles, enough to robustly identify gravitationally bound subhalos in the \subfind{} and \sublink{} catalogs.
For each subhalo in the FoF group that passes the minimum subhalo mass criteria, we utilize the \sublink{} catalogs to find the peak halo mass of each subhalo \citep{RG2015}.

As in \chambe{}, stellar masses are assigned to each subhalo in the FoF group using the abundance matching prescription of \citet{Moster2013}. 
Utilizing abundance matching to assign stellar masses allows us to circumvent simulation-specific stellar mass effects, and also results in a more straightforward process for applying our methodology to observational studies (see Sec.~\ref{sec:disc-suggestions} for more details). 
The peak halo mass and current redshift are used to calculate the stellar mass of each subhalo via the abundance matching prescription  $\ms{}=f(\Mpeak,z)$, where $z$ is the redshift of consideration.

@@

\subsection{Number of Pairs}\label{sec:pairprops-num}
We calculate the number of pairs by selecting all pairs at a given redshift that have not merged.
A single, non-merging pair with first infall at $z=3$ will contribute to the number of pairs at all redshifts from $z=0-3$.

@@

\subsection{Merger Fraction}\label{sec:pairprops-frac}
We calculate the merger fraction by dividing the number of pairs that merge (dashed lines in Fig.~\ref{fig:numorbits}) by the number of total unique orbits (solid lines) at a given redshift, including all pairs that do and do not merge. 
As before, an orbit with first infall at $z=2$ and merger at $z=1$ will be included in the merger fraction calculation for redshifts $z=1-2$.
Note, this definition of the merger fraction differs from that typically used in observational studies, where merger fractions are computed from close pair fractions with a correction term.

@@

% difference/similiarity between low and high
The merger fraction of low-mass and high-mass pairs is the same from $z=0-0.5$ and $z>2.5$. 
Between $z=0.5-2.5$, the high-mass merger fraction is higher than the low-mass merger fraction, and the knee of the merger fraction occurs at a lower redshift. 
We discuss this difference in more detail in Sec.~\ref{sec:disc-nonmergers}.
% This is likely because the number of low-mass pairs continues to increase before turning over and beginning to decrease at $z\sim2$, while the number of high-mass pairs continues increasing to $z\sim1$. 
% When the number of pairs stops increasing and beings to decrease, pairs that will not have time to merge before $z=0$ are being added, while many merging pairs are being removed from the sample, thus decreasing the merger fraction. 
% Since the turnover for the number of high-mass pairs occurs at a lower redshift, the knee of the merger fraction is likewise shifted to lower redshifts.
% move the above to discussion and end with something shorter: 
% This COULD be that the timescales. This will answer this with this study and discuss this in detail in Disc.~\ref. 
% could this be related to the number of non-mergers? is there a larger fraction of those starting at z=2 for the low mass galaxies?  these wouldn't be accounted for in Figure 6 right?
% So fig 6 shows that this isn't due to the merger time being significantly different. if anything the higher mass pairs merge longer.  you can state that at face value this difference in the knee may point to a difference in merger timescales, which you will illustrate in this paper is not the culprit.

@@

\section{Results: The Mass and Redshift Dependence of Merger Timescales}\label{sec:results}
    Using our sample of isolated low-mass and high-mass pairs in the TNG100 simulation, we calculate the merger timescales, or time until merger, for all of the merging pairs in our catalog. 
    The merger timescale of a pair is defined to be the amount of time that elapses in the simulation between the redshift at which pairs are selected and the merger redshift.\footnote{Since our orbits are only defined after infall, the merger timescales will likewise only be calculated for pairs that are post-infall at the given separation.}\textsuperscript{,}\footnote{We choose not to interpolate the orbits to get a more precise determination of the merger timescale, as the spread on the merger timescales in our study will be dominated by the variation of orbital configurations, rather than the uncertainty due to snapshot spacing. However,~\citet{Patton2024} recently showed that for studies needing more precise determination of, e.g., pericenter passages or merger times, their 6D interpolation scheme can improve timing uncertainties from $\pm80\Myr$ to $\pm3.3\Myr$.}
    Only the merging pairs will be considered for the remainder of the analysis.

@@

\subsection{Separation Dependence of Merger Timescales}\label{sec:results-timevsep}
    % how we calculate for this section
    We calculate the time until merger as a function of separation for low-mass and high-mass pairs at a variety of redshifts from $z=0.5-6$. 
    Binning the pairs by separation, we can study how the merger timescale changes for pairs selected at different points in their orbits.

@@

    We find that the time until merger for low-mass and high-mass pairs is positively correlated with the separation of the pair at each redshift. 
    % However, the slope of this correlation decreases at higher redshifts. 
    For the same physical separation, the merger timescale of high-mass pairs is shorter than for low-mass. 
    Pairs at $z=4$ with separations between $50-100\,\kpc$ tend to merge in less time than pairs with the same separations at $z=0.5$.

    In addition, the spread of times until merger and separations is smaller at higher redshift, for both low- and high-mass pairs. 
    The spread of merger timescales for low-mass pairs at $z=1$ goes from $0-8\,\Gyr$, while pairs at $z=4$ have merger timescales between $0-3\,\Gyr$. 
    Likewise, the spread of separations at $z=1$ is $0-200\kpc$, but at $z=4$ the spread is smaller, between $10-125\kpc$. 
    The high-mass pairs also have a larger spread in the distribution of merger timescales and separations at low redshift than at higher redshift.
    This is due to the growth of halos over time, and thus an increasing virial radius for FoF groups at lower redshifts.

@@

    \begin{figure*}[htb]
        \centering
        \includegraphics[width=\textwidth]{plots/bet-on-it/3_time_til_merger.png}
        \caption{(Top) The median time until merger as a function of redshift for low-mass (green) and high-mass (pink) pairs with 3D physical separations greater than $10\,\kpc$ and less than $(50,70,100,150,200,\mbox{and }300)\,\kpc$ from left to right. 
        (Bottom) The median time until merger as a function of redshift for pairs with physical separations greater than $10\,\kpc$ and less than $(0.25, 0.5, 1, 1.5, 2,\mbox{and }2.5)\,\Rvir$ from left to right. 
        % 
        The time until merger increases from $z=4$ to $z\sim0.5-0.75$, at which point the time until merger decreases to zero since all low redshift mergers must have short merger timescales to merge before the end of the simulation at $z=0$. 
        \kc{add h(z) line}
        }
        \label{fig:timescales-sep}
    \end{figure*}

@@

        % redshift evolution
        The median time until merger is $\sim0.5-0.7\,\Gyr$ at $z\sim6$, then rises to a peak of $2.3-2.4\,\Gyr$ at $z\sim0.6$, then decreases to 0 at $z=0$.
        The abrupt decrease is artificial rather than a true physical feature. 
        As $z\to0$, there is an increasingly small fraction of pairs that merge before $z=0$, as seen by the dashed lines in Fig~\ref{fig:numorbits}. 
        % These mergers must then proceed on timescales less than the elapsed time difference between a given redshift and $z=0$. 
        These mergers must also proceed on shorter timescales by definition, as they were selected as pairs that must merge prior to $z=0$ (i.e., a pair at $z=0.1$ that merges by $z=0$ can have a maximal merger timescale of $1.47\Gyr$, whereas a pair at $z=1$ can have a maximal merger timescale of $8\Gyr$). 
        This is shown by the dotted black line on the left of Fig.~\ref{fig:timescales}, which shows the lookback time of the simulation as a function of redshift.\footnote{The lookback time at a given redshift $z_n$ is equivalent to the time elapsed between $z_n$ and $z=0$. Thus, merging pairs cannot have merger timescales larger than the lookback time of the simulation at a given redshift.}

        In principle, merger timescales\footnote{Specifically, the elapsed time between crossing into the virial radius of the primary subhalo and coalescence with the primary.} for a body moving through a homogeneous field of collision-less matter can be analytically derived from the Chandrasekhar formula for dynamical friction~\citep{Binney2008}. 
        Departures from the idealized case introduce perturbations to that solution, the validity of which has been tested in cosmological hydrodynamic and N-body simulations~\citep{Jiang2008, BoylanKolchin2008}.

@@

        In Fig.~\ref{fig:timescales}, the black dashed line shows the Hubble time $1/H(z)$ multiplied by 0.25. 
        We did not perform a fit for this multiplicative constant, since we are interested in investigating the behavior of the merger timescale as a function of redshift rather than the overall scaling. 
        The redshift evolution of our median merger timescales is consistent with the scaling with $1/H(z)$ between $z=1.5-6$.\footnote{$H(z)$ is calculated using the same cosmology as the TNG100 simulation. Specifically, $\Omega_M=0.31$ and $\Omega_{\Lambda}=0.69$.}

        It is notable that such a simple scaling holds for both low-mass and high-mass pairs, especially in a complex simulation like TNG100 where subhalos are far from isotropic spherical distributions of mass, and that low-mass and high-mass pairs have the same merger timescales. 
        We believe this is due to the dominance of dynamical friction in determining the redshift behavior of merger timescales. 
        % Thus, the primary mechanics that determine merger timescales are likely dynamical friction

        % maybe discuss Poulton 2023 somewhere in here -- add after send to prelim committee

    \subsubsection{Physical Separation Selected Pairs }
    \label{sec:results-phys}
        Both separation selected samples have a lower separation criteria of $10\,\kpc$, to limit the impact of subhalos becoming indistinguishable in the \subfind{} catalogs.
        This lower separation criteria is also commonly applied to observationally selected pairs in studies of merger fractions and merger rates~\citep[see e.g.][and observational studies cited therein]{Lotz2011,Snyder2017,Besla2018,Chamberlain2024}.

@@

% ##################
% Discussion section
\section{Discussion} \label{sec:discussion}
    We have calculated the merger timescales, or time until merger, for a sample of physically co-located low-mass and high-mass pairs from $z=0-6$.
    In this section, we will explore the implications and broader impacts of our findings, and put our study in context with previous work.

    In Sec.~\ref{sec:disc-nonmergers}, we will discuss the non-merger portion of our orbit sample and the difference between low- and high-mass merger fractions.
    In Sec.~\ref{sec:disc-pairfrac}, we return to one of the key questions posed in our previous work: what is the underlying physical mechanisms that drives the difference in pair fraction evolution between low- and high-mass pairs. 
    Finally, in Sec.~\ref{sec:disc-suggestions}, we discuss the implications of our work for future observational and theoretical merger fraction and merger rate studies, and provide suggestions for a self-consistent framework moving forward.

    \subsection{Non-merger population}\label{sec:disc-nonmergers}
        In our catalog of 22,213 low-mass pairs and 3,029 high-mass pairs, only 16.66\% (3700) of low-mass pairs and 29.71\% (903) of high-mass pairs do not merge by $z=0$.
        In Sec.~\ref{sec:pairprops-frac}, we found that the merger fraction of low-mass and high-mass pairs at $z>3$ is $>0.95$. 
        The merger fraction decreases substantially as you move to lower redshifts. 
        While the decline in the merger fraction at $z\lesssim1$ is primarily due to the limited time that a pair will have to merge before the end of the simulation, we also find that between $z=\sim0.6-2.5$ the fraction of low-mass pairs is lower than that of high-mass pairs.

        One potential explanation for this behavior could be that low-mass pairs have shorter merger timescales than high-mass pairs, and thus merge at earlier times in the simulations, thus leading to a decrease in the merger fraction at higher redshifts. 
        However, through our analysis, we have shown that the merger timescales of our full low-mass and high-mass pair sample evolve in the same way from $z=0-6$. 
        This means that the difference in the merger fraction between $z=0.5-2.5$ is not a result of different merger timescales.

        We suspect that the cause of the difference between low- and high-mass merger fractions is not a difference in the merging population itself, but rather in the non-merger population and its evolution over time. 
        As we did not distinguish our non-merging sample into ``likely mergers" and ``fly-by" interactions for this study, we cannot say anything about how these two populations individually contribute to the difference in the merger fraction.
        However, one possible explanation could be that a larger fraction of infalling low-mass pairs at lower redshift results in fly-by interactions. 
        This could be the case if the velocity at which the secondary subhalo enters the primary's FoF group becomes increasingly large compared to the virial velocity\footnote{The virial velocity is the value of the circular velocity at the virial radius.}, resulting in fewer bound systems. 
        % We leave this as a focus of future work.

@@

    \subsection{Connections to Pair Fractions}
        In our previous work~\citep{Chamberlain2024}, we found that the pair fractions of isolated low-mass and high-mass pairs evolve very differently from $z=0-2.5$. 
        Low-mass major pair fractions are constant for $z>2.2$, but decrease by almost 70\% from $z=3$ to $z=0$ (see their Fig.~3). 
        High-mass major pair fractions slowly increase from $z=6$ to $z=0$, with a more significant increase from $z=0.25$ to $z=0$.

        One of the hypotheses we laid out in~\chambe{} for the difference in the behavior of the pair fractions is that perhaps low- mass and high-mass pairs have different merger timescales, which could lead to a difference in pair fractions from high to low redshift.

        However, in this study, we have shown that the merger timescales of low-mass and high-mass pairs are approximately equal at all redshifts, particularly when considering the full pair sample or the scaled separation selected sample. 
        This means that different merger timescales are not responsible for the difference in the evolution of pair fractions. 
        Rather, we suspect that the build-up of larger structures due to hierarchical formation under $\Lambda$CDM leads to a reduction in the total number of low-mass groups at lower redshift $z\lesssim2$.

        Additionally through this study, we have found that a very large fraction of the pairs selected in our previous work do indeed merge ($>90\%$ of pairs for $z>1.6$), so that the rate at which mergers occur is likely larger than the rate at which isolated low-mass pairs form, particularly at low redshift, thus leading to a quickly declining pair fraction measured in our previous work.

    \subsection{Implications and Suggestions for Future Observational and Theoretical Studies of Pairs}~\label{sec:disc-suggestions}
        While we are unable to make meaningful comparisons with previous observational studies of merger fractions and merger rates, as our study is completed using true physical separations from the simulation rather than projected separations, we can still draw meaningful conclusions about the application of our results to future studies of merger timescales and merger rates.

        \subsubsection{Pair Selection Criteria and Merger Rates in Illustris}
            Comparisons between studies of pair fractions, merger fractions, and merger rates, especially across cosmic time and different mass scales, are often challenged by the use of different sets of selection criteria in each study.

            On the observational side, pair selection criteria are often set by the observational parameters of the survey itself. 
            For example, limiting magnitudes and completeness limits can enforce a specific new range of stellar mass ratios to be considered (for example $1-1:6$ instead of $1:10$), while avoiding fiber collisions on a spectroscopic survey might lead to very specific separation criteria.
            On the theoretical side, it can be more straightforward to adopt a number of different pair selection criteria for comparisons to specific observational studies with identical criteria. 
            However, this means that theoretical studies typically restrict themselves to a subset of pair selection criteria specific to their target comparison studies.
